Early Detection of MCI that progresses to dementia - An Interdisciplinary Approach

### 2020-04-06 Source/Thesis/Thesis.tex
updated Thesis
update Thesis and Delphi Questionnaire
updated Pres

%%% PREAMBLE
\documentclass[10pt, letterpaper, twoside, openany]{book}

@@

\usepackage[utf8]{inputenc}
\usepackage{fdsymbol} 
\usepackage{graphicx}
\usepackage{float}

@@

\chapter{Introduction}
\include{Chapter1}

\chapter{Background and Related Work, Systematic Review of NLP and Machine Learning Research}
\include{Chapter2}

\chapter{Delphi Methodology and developing consensus on how best to collect language samples using technology}
\include{Chapter3}

\chapter{Development of a pipeline that processes language data accurately}
\include{Chapter4}

\chapter{Analysis of DementiaBank dataset}
\include{Chapter5}

\chapter{Pilot study of the methodology developed}
\include{Chapter6}

\chapter{General Discussion, Conclusions and Future Work}
\include{Chapter7}

>>>>

%%% PREAMBLE
\documentclass[10pt, letterpaper, twoside, openany]{thesis}

@@

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{float}

@@

\chapter{Introduction}
\section{Background}
Alzheimer's Disease (AD) and other forms of dementia affect a significant proportion of the geriatric population in the world today and is currently the sixth leading cause of death in the US and was named the leading killer of women in the UK. According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people. This represents 1 in 14 of those aged 65 or over and 1 in 79 of the general population \cite{AlzheimersSociety2014}. From a financial perspective, they estimate an annual spend of £4.3 billion of which approximately £85 million is spent solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually. Globally, this picture is a lot bleaker. Another report by Alzheimer's Disease International suggests that in 2015 there were 46 million people with a diagnosis of dementia globally and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}. The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars.
\par 
AD is a neurodegenerative disease in which a definitive diagnosis can only be produced at post-mortem. However, there are a number of psychological and physiological indicators that can indicate that dementia is present. From a physiological perspective, researchers have identified two proteins called beta-amyloid and tau. In a typical case, tau accumulates and eventually forms tangles inside neurons and beta-amyloid clumps into plaques which slowly builds up between neurons. At a certain point, the levels of beta-amyloid rise and trigger a more rapid spread of tau throughout the brain. Eventually, due to this and other changes, neurons lose their ability to communicate and the brain starts to shrink. This leads to some of the more psychological symptoms, those who have dementia demonstrate cognitive deficits such as problems with episodic and semantic memory, organizing and planning, difficulties with language, problems with executive function and visuospatial deficits \cite{McKhann2011}. In addition, these symptoms are often accompanied by emotional problems such as depression and behavioural difficulties. As more neurons die throughout the brain, a person with Alzheimer's gradually loses the ability to think, remember, make decisions and function independently.
\par
Despite this growing problem and an increasing understanding about how AD affects the brain there are no medications that improve the prognosis of those with AD. All the medications that are currently on the market are designed to manage symptoms. Whilst there are numerous investigational drugs in development for the treatment of AD, a larger than normal percentage (99.6\%) of these drugs fail in clinical trials (in contrast to anti-cancer drugs which have a 80\% failure rate) \cite{Cummings2014}. Researchers have proposed that a possible reason for the lack of success is that the drugs treatments are initiated too far along in the progression of the disease and thus much of the degeneration of the brain has already taken place \cite{Cummings2014}. 
\par
We can characterise a person's progression through AD in three stages. Whilst these stages are called different things in the literature, we will stick to the following naming convention throughout this thesis
\begin{enumerate}
	\item[a)] Stage 1: Preclinical AD
	\item[b)] Stage 2: Mild Cognitive Impairment due to AD
	\item[c)] Stage 3: Dementia due to AD
\end{enumerate}
Research has started to be more focused on stage 2 however a problem with focusing here is there are no clear defined boundaries between these three stages.
\par
One of the challenges of this approach is differentiating natural cognitive decline due to aging with decline due to a form of cognitive impairment or dementia. This challenge is often complicated further due to the large variation in the cognitive abilities and educational background of individuals. Albert and his team have worked to define clinical criteria which professionals can use to diagnose MCI due to AD and differentiate this from age-associated memory impairment and age-associated cognitive decline.  One of the most important observations from this piece of work is that a diagnosis of MCI requires evidence of intra-individual change and optimally requires evaluation at two or more points \cite{Albert2011}, and this is essentially to place more importance on the trajectory of a person's cognitive abilities rather than a person's cognitive ability in general. The criteria for MCI is detailed below \cite{Albert2011}.
\begin{enumerate}
	\item Concern regarding a change in cognition - A person or an informant should express concern that there is a change in cognitive ability in comparison to a previous level of performance.
	\item Impairment in one or more cognitive domains - There should be evidence of lower performance in one or more cognitive domains beyond what would be expected of a person given their age and education. 
	\item Preservation of independence in functional abilities - Whilst persons with MCI are expected to be able to maintain independence, it is common to experience mild problems in complex functional tasks which they may have been able to perform previously. This might mean that they take more time or be less efficient at completing these tasks, or it may be that may make more mistakes.
	\item Not demented - The deterioration should be mild to the point that there is no significant loss of functioning in social or occupational contexts.
\end{enumerate}
In addition to meeting the above criteria, a clinician must rule out other conditions or factors that could account for the decline in cognition with the goal to increase the likelihood that the underlying cause of this decline is dementia. 
\par
Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can, in some cases, slow the progress of the disease. Early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.
% Do more on the state of the art in terms of the analysis of biomarkers for MCI.
\section{Statement of the Problem}
Studies that explore ways in which we can diagnose MCI / AD generally follow one of two main approaches, the analysis of biomarkers such as concentrations of amyloid-β 1-42 (Aβ42) in cerebrospinal fluid (CSF) and analysis of cognitive abilities. The first approach yields reliable results in the detection of AD in its moderate and advanced states but does not perform well during the early stages of the disease. The second approach, that of analysing the cognitive abilities of patients in memory clinics, has gained more attention in recent years due to the fact that in clinical practice it has shown promise in the early detection of AD. In addition, the analysis of the decline of cognitive abilities is comparatively inexpensive and less invasive that the first approach which commonly requires the collection of a sample of cerebro-spinal fluid which is painful for the patient involved. This has a number of benefits for countries with less developed healthcare systems, or where the burden of healthcare is more extreme. 
\par
One of the most common ways in which clinicians traditionally use the analysis of cognitive abilities to make an early diagnosis of dementia is through the use of the Mini Mental State Examination (MMSE) \cite{Folstein1975}. The MMSE is a brief questionnaire consisting of eleven questions which tests cognitive aspects of mental function and requires only 5-10 minutes to administer \cite{Folstein1975}. The MMSE is chosen due to it's effectiveness at assessing a person's cognitive mental state at a specific point in time, as well as being as sensitive to changes as a more detailed and complex assessment such as the Wechler Adult Intelligence Scale \cite{Folstein1975}. Whilst the MMSE is useful as a brief screening tool it has it's limitations. The MMSE was not specifically created to screen for dementias and therefore does not interrogate key aspects of cognitive impairment known to be affected in dementia. It also has limited value in assessing under-educated subjects and a meta-analysis on the effectiveness of the MMSE as a diagnostic tool for dementia showed that it's accuracy was low (sensitivity between 78.4\% and 85.1\% and specificity between 81.3\% and 87.8\%). As the MMSE has been shown to have low accuracy specifically in the diagnosis of dementia, it becomes necessary for professionals to employ the use of other tools or measures such as the Free Cued Selective Reminding Test (FCSRT) \cite{Grober2010} or the Montreal Cognitive Assessment (MoCA) \cite{Davis2015}.
\par 
These tests have the benefits of being much more accurate at diagnosing cognitive impairment and discriminating between dementia and other types of cognitive impairment at the cost time and training of psychological professionals such as clinical psychologists in administering these tests. However, the utility of diagnosing dementia at the point where clinical intervention is warranted is limited because at this stage both psychological and pharmacological interventions have been shown to not be effective \cite{Prince2015, Cummings2014}. In order to further our understanding of the progression of dementia it is important to detect the signs of dementia before they are clinically apparent.  
\par 
Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent. Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time and alter cognitive function until a threshold is reached and clinical symptoms become more obvious \cite{Nestor2006}. The case of Iris Murdoch, who had a confirmed diagnosis of dementia, illustrates this theory well. Le et al \cite{Le2011} found, in their analysis of three writers and the novels they wrote, that Iris Murdoch's work declined subtly over time, but there was a steep drop off in the use of language in her last novel when, it is theorized, the symptoms of AD manifested themselves more significantly. If this theory holds true more generally, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed.
\par
The two main ways in which diagnosis is performed is through assessment of memory and language. Tests of memory are classically among the most accurate ways of diagnosing dementia, however these tests suffer from the same reliance on clinicians to administer these tests in a clinical setting. Language however is a lot easier to collect and can be done in more naturalistic settings. As with memory, these tests can be done over time and would be able to chart a patients language degeneration over time. Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on measure decline in the use of language in those with AD. There are a number of difficulties to watch out for with this approach. There are a wide number of factors that are involved in language degeneration in the elderly, and consequently there will be an expected amount of variability between subjects. The administration of such tests may induce nervousness and discomfort which may impact performance, and also repeatedly administering the same language tests for differences over time may be confounded by improved performance at tasks via practice effects. However, there is enough promise in this approach such that it could help further our understanding of the disease, it's progression and the parts of the brain affected in the early stages.
\par
McKhann and his team were tasked with developing diagnostic guidelines for dementia in such as way that the criteria were flexible enough to be used by general healthcare providers without access to specialist medical equipment.the DSM 5 \cite{McKhann2011}. They state those suffering with mild dementia generally encounter impaired language functions (speaking, reading, writing)––symptoms include: difficulty thinking of common words while speaking, hesitations ;speech, spelling, and writing errors. More specifically, they may substitute general terms for more specific terms and may avoid the use of specific names of acquaintances. There may be grammatical errors involving subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc. Those who have progressed from Mild to Major depression also have difficulties with expressive or receptive language. They will often use general-use phrases such as "that thing" and "you know what I mean" and prefers general pronouns rather than names. With severe impairment, sufferers may not even recall names of closer friends and family. Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur. Echolalia (meaningless repetition of another person's spoken words) and automatic speech typically precede mutism. With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.
\par
One of the most famous pieces of research on the topic of language decline in dementia was by Berisha and Liss who examined speeches and public interviews of former US president Ronald Reagan \cite{Berisha2015}. They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease. It turned out later to be the case that he had Alzheimer's Disease. Another classical study by Snowdon et al looked at whether linguistic ability in early life was associated with cognitive function and AD in later life \cite{Snowdon1996} . They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life. They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density in early life. As we can see, just with these two pieces of research the range of language deficits in those who suffer with AD are extremely variable and can differ from patient to patient as the disease progresses. The consensus among researchers that this language degeneration is typically accelerated by the presence of dementia \cite{Berisha2015} and that a potential indicator of dementia is the rate of change in which the decline occurs relative to a fixed point in time rather than a comparison across a cohort of individuals. 
\par
Emery \cite{Emery2000} completed a literature review looking at all the potential language deficits that could exist in those with AD and / or MCI. She divided these deficits into four levels of language: Phonology, Morphology, Syntax and Semantics. She proposed that language and the processes involved in language are hierarchical in nature and that language moves from simple units of construction (Phonology and Morphology), and build layers of complexity and sophistication (Syntax and Semantics). She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics. She asserted that the language forms we learn last are the first to deteriorate as we generally learn language in small simple units initially and build syntax and complexity as we are more comfortable with language. However it is important to note that different variants of dementia show different deficits in terms of language productions. Regardless, dividing language in this way is useful as it allows us to detect deteriorations in different parts of language usage and therefore may provide a way of discriminating between different forms of dementia. 
\par
It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD. However, whilst there is a move towards research aimed at looking more specifically at MCI we currently lack the measures that are sensitive enough to detect MCI. Given that we know language is affected before a clinical diagnosis of dementia is usually made \cite{Berisha2015, Snowdon1996, Le2011}, it makes sense to explore whether language on it's own can provide markers that may indicate a cognitive impairment that could progress to dementia. The field of machine learning and natural language processing has been suggested as a way to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline \cite{Boschi2017}. I theorize that we may be able to enable an earlier diagnosis of those with MCI and AD using samples of spontaneous speech, natural language processing (NLP) and machine learning (ML).
\par
There is a large body of research that looks at the decline in language in those with MCI and AD \cite{Taler2008, Boschi2017}. However there is conflicting evidence in these studies about which declining language factors are associated of MCI and AD \cite{Taler2008, Boschi2017}. Research therefore should look at these features in more detail and a clarification of this currently disorganised picture should go some way to helping researchers further understand the disease and it's progression. Another area of focus for research of this nature is the process of collecting appropriate language samples. Whilst collecting samples of language is comparatively unintrusive, researchers recognise that these samples require a rich sample of language that potentially cannot be generated by tasks such as the picture description task. Therefore, it would be useful to explore whether spontaneous discourse such a semi-structured interview, has the ability to put pressure on both the cognitive and linguistic systems in the same way as traditional cognitive tests such that it might be able to distinguish between healthy controls, those with MCI and those with AD. As is the case with other quantitative measures of cognitive ability, contrasting individuals with MCI and Early AD is challenging due to the variation in an individual's baseline speech capacity. It is therefore prudent to measure cognitive decline over time as suggested by Albert et al \cite{Albert2011}. There is some evidence to support this approach, Berisha et al \cite{Berisha2015} demonstrated through a longitudinal language analysis of spontaneous speech that there are marked differences in this process between those who would go on to have a diagnosis of AD and a healthy control. 
\par
\section{Research Questions}
The overall objective of the research described in this thesis is to investigate the deterioration of language in people diagnosed with MCI and the early stages of AD. In order to do this I aim to answer the following research questions
\begin{enumerate}
	\item What research has been conducted in this field so far?
	\item How can we use technology to best enable the collection of language samples?
	\item Can we develop a complete data processing pipeline that processes data efficiently
	\item What can machine learning tell us about how language affects us?
\end{enumerate} 
\par 
This thesis' original contributions are an increased understanding of how language deteriorates over time in those with MCI and early AD, the exploration via a delphi method of ideas for the collection of language features via smartphone or other internet of things devices and the development and execution of a pilot study to test the feasibility of collecting high quality language samples via in-home technology. I also contribute to the research into which language features are important in the diagnosis of MCI and early AD over time. Finally, I explore the potential of one class classification as a new approach to classifying those with MCI and Early AD vs controls. The potential impact of this research in this area is immense. 
\par 
A number of research themes combine both analysis of language and analysis of the acoustic component of speech. Whilst we acknowledge the existence of this large body of research and it's potential contribution to this problem, we have chosen to focus and attempt to understand the language analysis portion of this problem.

\section{Structure of thesis}
There is a lot research in this domain from a psychological perspective. In chapter 2, I look at the background and related work in the domain of language deterioration in those with MCI and AD from the psychological perspective. This includes looking at the current ways in which psychologists collect and analyse language samples and exploration of the features of language that psychologists feel are important in this area. In addition, I take a systematic look at the work of the experts in the fields of machine learning and natural language processing in the growing area of research for NLP and Machine Learning and it's application to this problem. 
\par 
The start of my research aims to answer questions around how we can use technology to aid us in collecting language samples that may facilitate the process of diagnosis. Currently, language samples are collected in memory clinics and physician's offices and is quite a time consuming process. More recently, work has been done to automate this process via a web based app called Talk2Me \cite{Komeili2019}, which seeks to use different psychological tests to assess cognitive decline through analysis of language. There are some potential flaws in this method of data collection.  In other areas, the use of smartphone technology has allowed the collection of language for the diagnosis of disorders such as Parkinson's disease and so it appears plausible to be able to use this technology and potentially other technologies such as Amazon Echo to facilitate the collection of language samples. The work described in chapter 3 therefore describes an exploration of experts opinions' on the following questions.
\begin{enumerate}
	\item How should we collect these language samples - Smartphones? IoT Devices?
	\item What tasks should we use to elicit these language samples
	\item At what frequency should we collect these language samples at
\end{enumerate}
\par 
An adapted Delphi survey methodology is used to develop consensus on the answers to these questions and the results of these are documented in this chapter. The resulting pilot study is described in Chapter 6.
\par 
Once we have a way to collect language samples, the question to consider is how we may best analyse this data. There are a number of factors that one must consider when answering the question. We look at the effectiveness of automatic speech recognition at translating speech into language, as well as the minimum quality of data that needs to be passed to any pipepline we develop. The Talk2Me app collects and then generates approximately 2000 lexico-syntactic, acoustic and semantic features {\cite Komeili2019} as part of their data collection framework. Chapter 4 discusses the challenges connected to constructing an effective data pipeline and contains two experiments.
\par 
Whilst Chapter 4 is more theoretical in it's content, Chapter 5 documents the first experiments that use this pipeline on three existing datasets. These datasets are, the Presidents Press Conferences, The Novels of three authors and the DementiaBank dataset.
\par
In chapter 6 Pilot study of the methodology developed
\par 
Finally, in chapter 7. I discuss conduct a general discussion of the results of my research. I also think about the strengths and weaknesses of my work and suggest ways in which the research have been improved. Lastly, I look at a number of areas for future work which can build on this research.

\chapter{Background and Related Work, Systematic Review of NLP and Machine Learning Research}
\section{Existing Neuropsychological Measures of Cognitive Impairment, and Repeatable Battery of Neuropsychological Tests}
In terms of standardised cognitive tests there are two main aims. Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility). This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their performance in terms of both diagnostic and prognostic utility.
\par 
\subsection{Repeatable Battery for the Assessment of Neuropsychological Status}
The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specifically looking at detecting a characterizing very mild dementia \cite{Randolph1998}. The authors felt that there was a shortfall of appropriate measures that were sensitive enough to milder impairments as well as a number of other shortcomings of existing tests. They met a number of design goals for this new battery of tests that addressed these shortcomings. The RBANS consists of a number of sub-tests across five distinct domains.

\begin{enumerate}
	\item Immediate Memory 
	\begin{enumerate}
		\item{List Learning}
		\item{Story Memory}
	\end{enumerate}
	\item Visuospatial / Constructional
	\begin{enumerate}
		\item{Figure Copy}
		\item{Line Orientation}
	\end{enumerate}
	\item Language
	\begin{enumerate}
		\item{Picture Naming}
		\item{Semantic Fluency}
	\end{enumerate}
	\item Attention
	\begin{enumerate}
		\item{Digit Span}
		\item{Digit Coding}
	\end{enumerate}
	\item Delayed Memory
	\begin{enumerate}
		\item{List Recall}
		\item{List Recognitions}
		\item{Story Memory}
		\item{Figure Recall}
	\end{enumerate}
\end{enumerate}

However, whilst the authors claim that the RBANS is adequately sensitive in person's with MCI \cite{Randolph1998} and there is some research to support this assertion \cite{Karantzoulis2013}, other research points out that the RBANS has poor sensitivity in detecting MCI \cite{Duff2010}. Another drawback of this battery is the lack of executive function measures and object naming tasks. Research has shown that the RBANS has good test-retest reliability and convergent validity. However, Duff et al warned that caution should be exercised when using the RBANS in a MCI population as it has lower sensitivity in this population \cite{Duff2010}.

\subsection{Digit Span Test}
The Digit Span Test (DST) is predominantly used to measure a person's working memory capacity, specifically the capacity used to store and recall numbers. A participant is presented with a series of numbers of fixed length and is asked to recall those numbers in normal or reverse order. The series of numbers gets progressively longer until such time as a participant fails three or more times out of eight presentations.
\par
Research has consistently shown that MCI patients have a significantly lower digit span score in both normal and reverse order versions of this test and a study by Muangpaisan has shown that the reverse order version of this test can, to some degree, predict a diagnosis of MCI \cite{Muangpaisan2008}. Muangpaisan also found that Age, Gender and Education have an impact on the performance of the tests. Emrani et al found that the DST revealed no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups. Notably, there was an attenuated recency effect in those with mixed MCI. \cite{Emrani2018}. Kessels identifies that that there are working memory deficits in MCI patients and these worsen with AD patients. \cite{Kessels2011} He also identified that both MCI and AD have impaired performance on all three conditions of the digit span test. No differences were found between forward and backward conditions in any of the groups. However, available tests may not detect subtle impairments \cite{Kessels2015}.
\subsection{Rey Auditory-Verbal learning test}
Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processes including short-term auditory-verbal memory, learning and retention of information. Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat. Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then repeat it again after 30 minutes \cite{Schmidt1996}.
\par
Several studies have shown that an impairment in RAVLT score reflect well the underlying pathology caused by AD. Thus making the RAVLT an effective early marker to detect AD in persons with memory complaints. Moradi investigated to what extent the RAVLT scores are predictable based on MRI data using machine learning approaches, as well as to find out what the most important brain regions are for the estimation of RAVLT scores \cite{Moradi2017}. They found a highly significant cross validated correlation between the estimated and observed RAVLT immediate and RAVLT Percent Forgetting. Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuracy comparible to MRI-based biomarkers \cite{Moradi2017}. Another study by Schoenberg found that RAVLT to best distinguish patients suspected of Alzheimer's disease from the psychiatric comparison group \cite{Schoenberg2006}.
\subsection{Digit Symbol Substitution Test}
The Digit Symbol Substitution Test (DSST) involves a key consisting of the numbers 1-9 and a corresponding unique symbol. Below this key is a series of numbers from 1-9 in a randomized order and repeated multiple times. The participant is asked to fill in the corresponding symbol for each number. The task requires that the participant move between the key and the randomized sequence such that  they may retrieve the correct answer from the key, hold this in short-term memory and transcribe the key in the appropriate place.
\par 
Among those with no disorder in cognition, mobility and mood, being in the lowest DSST quartile compared to the highest was associated with nearly twice the odds of developing one or more clinical or subclinical disorders. Associations were stronger for incident clinical disorders in cognition. Slower psychomotor speed may serve as a biomarker of risk of clinical disorders, mobility and mood. While in part attributable to vascular brain disease, other potentially modifiable contributors may be present \cite{Rosano2016}. Further studying the causes of psychomotor slowing with ageing might provide novel insights into age-related brain disorders. Pascoe compared patients with PD with Normal Cognition (PD-N) with those with PD and MCI (PD-MCI) and healthy participants. PD-MCI participants achieved significantly lower scores than other groups in the DSST task \cite{Pascoe2018}.
\subsection{Verbal Fluency}
There are a number of tests that characterise this category of verbal fluency, but generally fall into two categories. Letter (Phonemic) fluency involves the generation of as many words as possible which begin with a specified letter. Category fluency involves the generation of as many words as possible that fall into a specfied category. Both these tasks impose demands on a number of different cognitive processes namely, executive function, verbal retrieval and recall, giving appropriate answers while monitoring previous answers and inhibit inappropriate responses. However, these two tasks require different strategies when attempting them. Letter fluency relies on search strategies based on lexical representations whereas category fluency requires a search for semantic extensions of a superordinate term, meaning that semantic associations within the lexicon must be intact in order for the task to be carried out successfully. 
\par  
There have been numerous studies which have documented the impact that AD has on verbal fluency tests in both categories. A review carried out by Henry et al \cite{Henry2004}, found that performance in both letter fluency and category fluency was impaired in those with AD vs controls, but found a larger effect for tests of semantic fluency.
\subsection{Naming Tests}
Word-finding difficulty is a common symptom of AD and these deficits usually occur during the early stages of the disease progression. As such , a test of a patients ability to find words (known as confrontation naming) is a common way to measure cognitive decline. One common way to do this is the Boston Naming Test (BNT) which comprises 60 items on a spectrum of very frequent to very infrequent. This has been reduced subsequently to two thirty item versions and four fifteen item versions, which correlate significantly with the original sixty item version and the benefit of the shorter versions of the test is that it facilitates testing of individuals with AD who may suffer from fatigue or limited attention span \cite{Williams1989}.
\par 
Willers et al \cite{Willers2008} studied Twenty aMCI patients, twenty AD and 21 controls matched by age, sex and education level. They found AD patients obtained significantly lower total scores on the BNT than aMCI patients and controls. aMCI patients did not obtain significant differences in total scores but showed significantly higher semantic errors compared to controls. Semantic processing is impaired during confrontation naming in aMCI.
\par 
Vadikolias investigated the impact of education on naming tests \cite{Vadikolias2012} and found that higher educational attainment in aMCI subjects were correlated with better performance in verbal and non-verbal tasks during repeated examinations over 1-year. Subjects with a lower level of education performed worse than patients with a high level of education who presented a more stable clinical score. The explanation for this is the idea of a 'cognitive reserve' in participants with a higher education such that this provides a buffer that, while not preventing the physiological symptoms of AD, can potentially delay the clinical onset of cognitive symptoms that characterise AD. This theory is supported by Snowdon who found a relationship between early life linguistic ability and the density of neurofibrillary tangles in his nun study \cite{Snowdon1996}. Whilst these studies and tests provide evidence that there are word finding difficulties in those with MCI, on their own they do not provide sufficient ability to diagnose MCI or provide an prognosis of disease progression.
\subsection{Rey-Osterrieth Complex Figure Task}
The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory. The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944) \cite{Rey1941, Osterrieth1944}, asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition). In the immediate copy condition, the complexity of the figure requires an integrative cognitive ability. The reproduction of such a complex structure involves processes such as planning and organizational strategies that are related to executive functions. In order to make this task repeatable, Taylor et al designed a comparable set of figures which have proven to be of equivalent difficulty \cite{Taylor1969, Strauss1990}. 
\par
Salvadori found that patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory \cite{Salvadori2018}. Evidence shows that patients with disorders that possibly involve attention and executive functions are characterized by a more disorganised approach when copying the ROCF compared to controls. One of the difficulties with the presentation of this task in a repeated battery will be the fact that it turns from an incidental memory task (it is a memory task but the participant is not forewarned that they will be required to memorise the picture) into an intentional memory task (given the previous exposure to the task, it would be expected that a participant would preempt the delayed recall portion of the test and spend more time attempting to memorise the details) \cite{Teng2009}. 
\subsection{Hayling Sentence Completion Test}
Inhibitory deficits are a common in all stages of dementia. This is usually tested using Stroop test, however this has the drawback of lacking ecological validity. Therefore researchers have moved towards using the Hayling Sentence Completion Test (HCST) which uses skills such as word retrieval as well as the ability to inhibit ones responses where appropriate and is therefore a much more ecologically valid task. There are two parts to the HCST. In the first part, participants have to complete a sentence by providing a word that best fits the given sentence (this is known as the initiation condition). In this second part, participants have to complete sentences by inhibiting an impulse to give the word that best first the sentence as in the first part, and producing a semantically unconnected word. Performance in both conditions is measured by the time taken for the participant to initiate a response, and in the inhibition condition also by the correctness of the word.   
\par 
Martyr et al compared healthy controls with patients with dementia and patients with Parkinson's disease \cite{Martyr2017}. They found that a high proportion of Category A errors (producing a word that fits the sentence when instructed otherwise) was a factor in performance loss for participants with dementia. Findings suggest that the HSCT may be sensitive to verbal suppression deficits and may provide insight into inhibitory control in participants with dementia. Patients with Dementia were significantly slower than controls in the initiation and inhibition conditions vs healthy controls, and slower than patients with Parkinson's disease in the initiation but not the inhibition condition.
\subsection{Grooved Pegboard Test}
The Grooved Pegboard Test (GPT) was originally designed to cover a variety of different psycho-motor functions including hand-eye coordination and motor speed. However, some studies have shown a correlation of performance in the GPT and measures of cognitive performance such as the Montreal Cognitive Assessment.
\par 
Bezdicek found that the GPT predicted performance on the MoCA and concluded that in addition to being a measure of motor skills, there were results that showed that the GPT could also provide information about a participants cognitive skills \cite{Bezdicek2014}. Whilst this was specifically using a population of patients with Parkinson's disease, given the MoCA and GPT have shown degree of correlation, there remains an opportunity for research that explores the use of the GPT with MCI and Early AD populations.
\par 
Darweesh et al, used the Purdue Pegboard Test (PPT) to assess manual dexterity in a healthy older adult population and followed their participants for between eight and twelve years until an indication of the onset of a neurodegenerative disease \cite{Darweesh2017}. In this time 227 (4\%) of their participants went on to develop a diagnosis of dementia. They noted that higher PPT scores were associated with lower risk of incident neurodegnerative disease and noted significant associations of PPT scores with all forms of dementia and this potentially highlights a deterioration of motor function in the pre-clinical phase of dementia.
\subsection{Visual Search}
The visual search task requires participants to be watching a computerised display in which a target appears either by itself or surrounded by other elements which are used as a distraction. These other elements can be vary in similarity to the target. This is inherently a measure of attention shifting and processing of various similarity and as one would expect, the visual search time is increased in the presence of distractors. This effect is magnified in participants who have a cognitive impairment. 
\par
Research has shown there to be deficits in both AD and MCI populations in this task \cite{Tales2005}, and those with MCI exhibited less severe deficits compabut that these deficits were not as apparent in MCI populations, there remained a significant enough difference that could be used to differentiate healthy controls from those with MCI. In a comparison of visual search performance between those with aMCI and healthy controls, Tales noted significantly poorer performance in the aMCI group. However, she also noted a good deal of heterogeneity in the aMCI group which illustrates that whilst the aMCI group have essentially the same condition, presentations within this population can differ markedly. The results from this study also illustrates the use of a non-memory task as a means of diagnosing dementia as patients who went on to be diagnosed with dementia whilst this study was in progress had significantly poorer visual search performance.  
\subsection{Free Cued Selective Reminding Test}
The Free Cued Selective Reminding Test (FCSRT) was borne out of the premise that by controlling the conditions of learning, a measure of memory is possible that is not confounded by normal age related changes in cognition. Theoretically speaking, any controlled learning test should be able to discriminate between cognitive decline due to age and cognitive decline due to a cognitive impairment. The FCSRT starts with a learning phase in which participants are asked to look at a card containing four pictures (e.g., grapes) for an item that belongs to a named category(e.g., fruit). After these four items are identified, immediate cued recall of these four items is tested and this is repeated for a total of 16 items. Following this learning phase is the recall phase in which participants are asked to recall all the items identified without cues and any items which are not recalled are then cued. There are three outcome measures for this test, free recall (the number of items recalled without cues), total recall (the number of items recalled with or without cues) and cue efficiency which is identified as follows \cite{Grober2010}.
\begin{equation} \label{x1}
CueEfficiency = (totalRecall-freeRecall)/48-freeRecall, range 0.0-1.0)
\end{equation}
Grober found that patients with impaired free recall with four times more likely to develop dementia than those with intact free recall \cite{Grober2010}. This test has also been shown to distinguish patients with MCI who then went on to develop AD, from those with MCI that did not then go on to develop AD and this led the researchers to used the test to categorise prodromal or the amnestic syndrome of the medial temporal lobe by FSCRT score \cite{Sarazin2007}. 
\subsection{Conclusions}
I have looked at a number of different cognitive tests and a battery of tests that aim to have high diagnostic and prognostic capabilities in the MCI population. However, particularly with the RBANS, there lacks sufficient sensitivity in differentiating those with MCI from healthy elderly individuals such that this tool could be used with a level of confidence in a clinical setting. In regard to the tests, a common theme is a lack of studies and therefore evidence into the utility of these tests with an MCI population. However if, as researcher, we aim to investigate this population further then a benchmark battery of tests which is sensitive enough in both diagnostic and/or prognostic utility should be a goal. 
\par 
Another criticism of the current literature is the lack of consistency with regard to the experimental groups. For example, some studies focus on differentiating between an MCI group, a AD group and a healthy controls group whereas other studies may further subdivide the MCI category according to a number of factors such as Amnesic or Non-Amnesic MCI, Mixed MCI (sometimes called Multi-domain MCI in the literature). Given the inconsistency in defining the experimental groups the confusing and often conflicting results that researchers produce is to be expected. Future research should look at the standardization of the operational definition of cognitive impairment in MCI may result in more consistent predictions of progression to AD. 
\par 
Finally, it can be said that in testing for MCI and comparing these results with a cohort of those with a similar diagnosis has limited utility. This is because there is huge variability in the presentation of MCI within participants that even if controlled for with a matched pairs design for age, gender and education will produce inconsistent results. There is an argument therefore for the use of longitudinal studies, where the comparisons on the performance of these tasks are with the participants themselves. There are not many studies which use a longitudinal approach, although those that do show promising results.

\section{Types of Language Assessment}
One of the key debates when looking at how to analyse language is the type of task provided to elicit language production in participants. In the literature researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data.
\par
\subsection{Picture Description Tasks}
One of the most commonly used tasks to measure language is the Picture Description task. An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Borod1980}. The Cookie Theft picture (pictured below) depicts a scene of a home typical of the period of time when it was created and would generally not require participants to use any complicated vocabulary to describe. In this task participants are asked to describe the picture presented to them in as much detail as possible.  This task was originally designed to assess Aphasia, but has shown itself to be useful in the assessment of language for the purposes of diagnosis of MCI and AD as well \cite{Giles1996}
\begin{figure}[H]
\centering
\includegraphics[width=240px, height=150px]{images/BCTPicture.png}
